\chapter*{Biography}
\addcontentsline{toc}{chapter}{Biography}


Ankit Kaushik (born on June 7, 1984) received a B. Tech degree in electronics and communication engineering from Guru Gobind Singh Indraprastha University, Delhi, India in 2005. He further received a dual M. Sc. degree in information and communication technology from University of Karlsruhe (now Karlsruhe Institute of Technology), Karlsruhe, Germany and Politecnico di Torino, Turin, Italy in 2007. From 2007-2012, he was with Leica Camera AG, Germany, where he worked as a Design Engineer. 

Since 2012, Ankit Kaushik has been with the Communications Engineering Lab, Karlsruhe Institute of Technology, as a Research Associate. During the winter semester 2015/2016, he was Visiting Researcher with the Security, Reliability and Trust (SnT), University of Luxembourg, Luxembourg. His research interests include software defined radio, cognitive radio communications and networks. He was a recipient of MERIT Scholarship for his Masters Studies within the Erasmus Mundus Scholarship Program and Subjective Winner of 5G Spectrum Challenge held at 2015 IEEE DySPAN conference.

%\section*{Personal Details}
%\begin{tabular}[]{ll}
%	Date of birth: & June 7th, 1984 \\
%	Place of birth: & New Delhi, India \\
%	Citizenship: & Indian
%\end{tabular}
%
%\section*{Education}
%
%\begin{tabular}[]{lp{6.95cm}}
%	
%	04/2012 - 09/2016: & Research Associate, \\
%										 & Karlsruhe Institute of Technology \\
%	09/2007 - 03/2012: & Design Engineer, Leica Camera AG, Solms, Germany \\
%	10/2005 - 07/2007: & M.Sc., MERIT Karlsruhe Institue of Technology, Germany and Politecnico di Torino, Italy \\
%	08/2001 - 06/2005: & B.Tech., Guru Gobind Singh Indraprastha University, Delhi, India \\
%	04/1987 - 06/2001: & Oxford Senior Secondary School, Delhi, India 	
%\end{tabular}

