\chapter*{Acknowledgements}

My time at Communications Engineering Lab (CEL) has been a wonderful experience. This would not have happened without the presence of some prominent personalities in my life. First and foremost, I would like to sincerely thank my supervisor Prof. Dr.\,rer.\,nat. Friedrich K. Jondral for accepting me as a Doctoral candidate. %Despite my lack of proficiency to the German language, he  
Despite my divergence to the industry, after my masters studies, he showed immense faith by giving me an opportunity to fulfill one of my goals. 
Besides, I admire him for his patience, guidance and providing me a suitable environment to carry out research. I would also like to thank Prof. Dr.\,-Ing. Anja Klein for agreeing to co-supervise my thesis.

My sincere thanks go to my colleagues at CEL, in particular, Dr.\,-Ing. Holger J\"akel, Dr.\,-Ing. Ralph Tanbourgi, Dr.\,-Ing. Noha El Gemayel and Dipl.\,-Ing Johannes Fink, who have been extremely supportive and always remained open for discussions. I would also like to thank Beate Mast and Brigette Single for handling the administration, Angelika Olbrich for helping me out with graphics, and Reiner Linnenkohl and Peter Herttrich for providing me the necessary infrastructure. During this time period, I had an opportunity to supervise students and I must agree to the fact that students are the best teachers. I am grateful for their contributions and extensive discussions of all forms. 

Outside CEL, I had an opportunity to collaborate with Dr. Shree Krishna Sharma, Prof. Symeon Chatzinotas, Prof. Bj\"orn Ottersten from University of Luxembourg. I am extremely thankful for their insightful suggestions that have been beneficial for my research. My special thanks to Dr. Sharma for showing interest in my research, proof-reading my thesis and conducting fruitful discussions. 

Last but not least, I am indebted to my parents Poonam and Rajesh, its their hard work, sacrifices and most importantly their belief that have brought me this far. In particular, my immense gratitude goes to my wife Mini for her love, support that kept me going during the stressful periods and giving me the most beautiful gift of my life, our daughter Alaiza. I must admit that I am extremely grateful to have such a life partner. 
